---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

The data is preprocessed such that the interval column is changed to be minutes into the day, meaning the value will be between 0 and 1435. Originally the interval is of the form "%H%M", so without preprocessing there would be jumps e.g. going from 155 to 200.


```{r}
if (!file.exists("activity.csv")) {
  unzip("activity.zip")
}
act <- read.csv("activity.csv", colClasses=c("numeric", "Date", "character"))
act$interval <- sapply(act[,3], function(x) {
  if (nchar(x) < 3) {
      as.integer(x)
  } else if (nchar(x) == 3) {
      as.integer(substr(x, 1, 1)) * 60 + as.integer(substr(x, 2, 3))
  } else {
      as.integer(substr(x, 1, 2)) * 60 + as.integer(substr(x, 3, 4))
  }
  })
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
steps <- tapply(act[,1], act[,2], sum)
qplot(steps, binwidth=1000)
mean(steps, na.rm=TRUE)
median(steps, na.rm=TRUE)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
